Clouds and Radiation in a mock-Walker Circulation

### 2019-10-14 Silvers_WalkerCell.tex
many changes to Silvers_WalkerCell including a new table and significant text changes.

% What are climate modelers doing to reduce uncertainties about cloud processes?
The representation of clouds in climate models is the most stubborn contributor to uncertainty of the climate response to increased 
greenhouse gases.  A few of the reasons clouds are difficult to represent in models are imprecise definitions of clouds, an existential 
dependence on trace amounts of condensed water, the importance of microphysical processes at scales less than the grid-scale, 
and the close coupling between atmospheric radiative fluxes, clouds, and the large-scale dynamical circulations of Earth.  To elucidate this coupling between clouds, radiation, and the large-scale circulations of the Earth, we simulate an idealized equatorial tropical Pacific.   In the western Pacific warm sea surface temperatures prevail beneath rising atmospheric motion and deep convective storms.   In contrast, the eastern tropical Pacific is characterized by cooler sea surface temperatures and low-level, non-precipitating clouds with a dry atmosphere above the clouds.  These regions are connected by a large-scale circulation.  We analyze this scenario in a high-resolution and coarse resolution simulation and infer how deep convective heating in the west Pacific can influence the atmospheric cooling thousands of kilometers away in the east Pacific.  This cooling is a key ingredient in the determination of the Earth?s sensitivity to changes in greenhouse gas concentrations.

@@

@@

@@

\begin{table}
\begin{center}
\caption{A few specifications of the simulations used most heavily in this study.  The precipitation values in parenthesis are for the experiments
in which LWCRE is off.}
    \begin{tabular}{*{6}{c}}
    \hline
    \hline
 Name & Grid Spacing $(\mathrm{km})$ & Domain $ (\mathrm{km^2}) $& Length $(\mathrm{months}) $ & Convection & Precip $(\mathrm{mm/day})$   \\ \hline
  P100 &  100          &   800 $\times$ 16000    &  60              & parameterized           &  4.1 (3.5)          \\ 
    \\
  P100 &  100                & 800 $\times$ 4000     & 60            & parameterized         & 3.9                \\  
    \\
  P25 &  25             & 200 $\times$ 16000      & 60             & parameterized          & 4.0 (3.8)                \\  
    \\
  P25  &  25             & 200 $\times$ 4000      & 60             & parameterized           & 3.84 (3.66)               \\  
    \\
 E25  &   25          & 200 $\times$ 4000      & 60             & explicit          & 3.74 (3.46)               \\  
    \\
 E2   &   2          & 100 $\times$ 4000      & 6             & explicit          & 3.11 (3.50)                \\ 
    \\
 E1   &   1          & 10 $\times$ 4000      & 6             & explicit          & 3.36 (3.74)                \\  \hline

@@

There are circulation cells below $500 \rm{hPa}$  and above $500 \rm{hPa}$.  Deep-convective activity dominates the 
region over the warm pool, shallow convection is common over a wider range of SSTs, and for the most part stratocumulus 
clouds are absent.  When the coupling between the circulation and clouds is broken, by turning off the longwave CRE, the atmospheric 
state is much more symmetric about the maximum SST.

We perform a series of sensitivity experiments to highlight the different ways in which these experiments can equilibrate.  To 
show the sensitive nature of the distribution of the precipitation maximum we test the impact of convective parameterization, 
longwave radiation interactions with clouds, domain size, 
and the resolution, or grid-spacing.    Muller and Held, 2012 found that self-aggregation was sensitive to both domain size and resolution
because of the sensitivity of the low-level clouds to the longwave radiative cooling.

The abnormality, or asymmetry, of the P25 simulation was the first surprising feature noticed by the authors.
This asymmetry is also present on larger domains and at a resolution of $100 \, \rm{km}$.  For some of the other cases observed, such
as P100 (on the small domain), the asymmetry can be described as an aversion of the precipitation maxima to reside over the 
SST maximum while in other cases, such as the P25 case shown in \ref{fig:rh_psi_P25vsE25}, a strong ($1 \, \rm{m s^-1}$) domain
mean shear develops which shifts the precipitation and circulation off center for years at a time.  This appears to be not entirely a 
result of the convective parameterization, but due to an interaction between the convective parameterization and the LW CRE.

@@

We find large differences among the experiments in the domain mean precipitation, condensate/clouds, and circulation.  In particular
the experiments with parameterized convection are much more erratic.  Above about 500 \textit{m} the P25 experiment
has strong domain mean wind shear that clearly influences the structure of the precipitation, as seen in 
five year Hovmoller diagrams.  The degree to which this shear influences the comparison with other experiments is unclear.

@@

@@

One of the most prominent features of our 25km grid-spacing simulations is an asymmetry in both the time-dependent and 
steady-state solutions.  This results in the steady-state precipitation maximum locating off the maximum in SST and is apparent in 
the vertical velocity, mass circulation, relative humidity, specific humidity, and radiative heating.   A tendency of the precipitation to 
occur off of the SST maximum is also apparent in the simulations with a grid-spacing of 100km, although the 25km simulations display 
an irregularity that is greater than the 100km simulations.  Complex patterns of precipitation over a fixed sinusoidal or Gaussian SST 
distribution has been noted many times 
(Bretherton et al. 2006, Kuang and Wofsy, Ming Zhao (personal communication, see Jeevanjee et al. 2018)) but the irregularities have 
tended to be symmetric, unlike our results.

For the simulations that are closest to the set up of a GCM, with fully parameterized convection, the steady-state circulation and 
precipitation maximum are located not over the warmest SST but shifted to regions about one degree cooler.  For the case
with a grid-spacing of 25km and a domain width of 4000km, the precipitation maximum can remain shifted off of the maximum in
SST for multiple years.   While focused within the central quarter of the domain, simulations on larger domains and with a 
grid-spacing of 100km also show an aversion to settling over the SST maximum.   When the convective parameterization
is turned off, the overturning circulation becomes weaker and broader (w, and psi), but the precipitation, cloud fields, and 
circulation all consistently reside over the SST maximum.  When the convective parameterization is turned on, a domain mean 
wind shear develops throughout the lowest 10km that acts to shift the convective activity and circulation off of the SST maximum.  
Note: check the wind shear for P25large, P100small, and P100large.

@@

\subsection{The Influence of Domain Size on the LWCRE and the LS Parameterized Precipitation}

@@

When designing numerical experiments the size of the domain and the grid-spacing that determines model resolution are two of the most 
fundamental choices that must be made.   One the goals of this study was to explore the results of changing grid-spacing over a wide 
range of values.  To simplify the analysis we have in most cases chosen to keep the 
long horizontal dimension fixed at $4000\, \rm{km}$.  However, previous studies (e.g. Bretherton et al. 2006) have 
documented sensitivities to domain size.   The analysis of the previous, and of the next section focus on results from experiments using
a domain with a long domain length of $4000 \, \rm{km}$.  However we do find sensitivity to the domain size and thought it to be
sufficiently interesting to report.

Examining the evolution in time of the precipitation field reveals important characteristics of these simulations, and 
illustrates how much the distribution can vary as a function of domain size, parameterization of convection, and the effect of the 
longwave radiation due to clouds. 
Shown in Figures {\ref{fig:domdep}} and {\ref{fig:domdep_lwoff}} are Hovmoller plots of precipitation after 
averaging along the short horizontal dimension.  The four panels show simulations with a grid-
spacing of $25\, \rm{km}$ on a grid with a horizontal width of $4000\, \rm{km}$ (far right) and $16000\, \rm{km}$ (middle right), and 
simulations with a grid-spacing of $100\, \rm{km}$ on a grid with a horizontal width of $4000\, \rm{km}$ (middle left) and 
$16000\, \rm{km}$ (far left).  Figure  {\ref{fig:domdep_lwoff}} shows the equivalent simulations with LW CRE off.    
 At all resolutions the Hovmoller plots show that the longwave CRE acts to concentrate the 
precipitation over a smaller geographic extent.   This is consistent with the idea that CREs strengthen an overturning circulation and
narrow the region of deep convection (see Popp and Silvers, 2017, HH, Dixit et al. 2019.  etc.).
The structure of the precipitation depends strongly on domain size and shows less dependence on grid-spacing.
On the large domains, the difference between experiments with and without LW CRE is extreme.  In contrast to the control 
experiments in Figure {\ref{fig:domdep}} which all show a narrow region of strong precipitation which meanders within 
500 km of the SST maximum at the center of the domain,  
the experiments without the LW CRE basically have no region with consistently strong precipitaiton.  Instead, 
smaller cells or lines of precipitation develop within an area that is roughly 8000 km wide.

@@

Smaller domain lengths were found to have a more focused ascent region and larger precipitation rates in Bretherton et al., 2006.
They also found less low-level clouds over the colder SSTs in the smaller domains, relative to a control domain size that was 4x
larger.   We have found larger precipitation rates for the larger domains.  Does our larger domain ascent region change?

@@

We hypothesize than an important difference between our simulations and previous studies which use GCMs
is that although the deep and shallow convection is parameterized
with the GFDL double-plume parameterization, the fraction of precipitation that is produced by the convective parameterization 
is much smaller in our simulations than it is for many GCM-type experiments with parameterized convection.   At 25km grid-spacing, 
the majority of condensational heating occurs through the large-scale cloud  scheme.   See Figure \ref{fig:conv_vs_ls}.
%Overall, the fraction of precipitation that is 
%due to the large-scale scheme is roughly 40\%, and in many places it is actually greater than the amount of precipitation 
%produced by the convection scheme.   
%In comparison, many GCMs, have between 70\%-90\% of their precipitation produced by 
%the convective parameterization.  
There are further surprises in the fraction of precipitation that is large-scale.  On our small 
domains, with a bowling alley domain length of 4000 km, the simulations with 100km grid-spacing have values of $f_{ls}$ 
of $75-80\%$.  Often, at GCM-like resolutions such as 100km the convective precipitation would be expected to control much of the
precipitation rather than the large-scale parameterization.  However, when the domain length is increased to 16000 km
$f_{ls}$ decreases to a range of $0-70\%$ that depends strongly on location.  Over the warmest region of the surface the 
precipitation is still dominated by the large-scale scheme.  However, turning off the longwave CRE leads to a dramatic reversal 
in which $60-75\%$ of the precipitation is convective over the warmest SST.  The large domain without the longwave CRE is the 
only configuration in which the convective precipitation is dominant.    If the default, or baseline, configuration of a GCM is built 
with the expectation that the convective parameterization scheme will dominate the production of tropical precipitation, then 
perhaps a configuration that minimizes the convective precipitation will result in a balance between the parameterization 
schemes that is less intuitive, such as an asymmetric response to symmetric forcing.

Below is a list of several changes that occur when the LW CRE is off.  The points which have not already been discussed will be later in the paper.

Turning the longwave cloud radiative interactions off tends to:

\begin{itemize}
  \item decrease the strength of the overturning circulation and the domain mean precipitation
  \item increase the midtropospheric RH by $5-15 \% $
  \item eliminate the difference of liquid condensate at 900 hPa between explicit and parameterized simulations.
  \item weaken the horizontally oriented low-level circulations.   This looks connected to a contracting of relative 
           humidity maxima around 600 hPa.
  \item spread out in geographic space the precipitation maxima
  \item dramatically homogenize the surface enthalpy flux over the warm pool, primarily through the latent heat flux
  \item decrease both the LTS and the EIS, over both the warm and cold regions
%  \item in most cases the symmetry about the SST maximum is increased.
  \item decrease the lowest level $q, \theta_e,$ temperature $T$ and virtual temperature $T_v$.  
  \item eliminate large differences in the evaporation and lowest level temperature in the center of the domain between 
           the 1 and 2 \textit{km} simulations. 
  \item increase evaporation in the 1 and 2 \textit{km} cases, decrease it in the 25 \textit{km} case.
  \item amplify the larger values of ice condensate aloft for the explicit simulations.  
\end{itemize}

@@

It is clear that the overturning circulation in the tropical Pacific encompasses dynamical motions at a host of scales and that the 
different cloud types interact with the circulation on scales ranging from meters to thousands of km's.  
The tropical Pacific also serves as an ideal location to study interactions between clouds and the 
circulation because it includes strong overturning circulations, abundant shallow cumulus, congestus, and cumulonimbus 
clouds (Johnson et al. 1999) as well as stratocumulus cloud decks along the eastern extremities of the basin.
Can we hope to simulate all of this in an idealized modeling context?    
Our initial motivation for using the GFDL AM4 model on this doubly-periodic domain was to simulate this region 
%that is rich in phenomena 
in the context of both a GCM and a CRM in the hopes of learning something from the high resolution 
cases that will help us to improve the development of the GCM.    
The dynamical core of the current generation GFDL models is ideally suited to this task because it can easily 
be used over a wide range of resolutions and can solve either the hydrostatic or nonhydrostatic 
governing equations.  In this section we focus on simulations with grid-spacing of 1km, 2km, and 25km.   
A glance at Figure \ref{fig:hov_4mods_6mn} shows notable differences in the precipitation structure that result from the 
overturning circulation at different resolutions.  Shown are 180 days precipitation from the P25, E25, E2, and E1 simulations.  As the 
resolution increases towards $1 \, \rm{km}$ the distribution of precipitation becomes more regular and consistently 
centered over the SST maxima in the middle of the domain.  Both the P25 and E25 simulations show more variability at later 
times, compared to these first 180 days.  The simulations with explicit convection at resolutions typical of cloud-resolving 
models (E2, E1) show little aversion to the precipitation maximum occurring over the maximum in SST and relative to 
the P100, P25 and E25 simulations 
they are able to maintain a smoother distribution of  precipitation values over a broader range of SST values.

@@

Because of the close connection between the domain mean value of precipitation, atmospheric condensational heating, and 
the total radiative cooling the time evolution of the precipitation serves as a useful measure of wether a model has 
reached statistical equilibrium.  Figure \ref{fig:precip_dom_mn} demonstrates that after about 30 days, the P25, E2, and E1
simulations begin to oscillate about a mean state that is roughly constant over periods greater than about 30 days. 
The large oscillations in domain mean precipitation are similar to those noted in previous studies 
(Silvers et al., 2016; Patrizio and Randall, 2019).   The experiments with a higher resolution have a lower value of domain
mean precipitation, with the E2 simulation having the lowest (3.1 mm/day).  This can be understood as a consequence
of the differences in upper level cloud fraction and will be further discussed.

Superposing the circulation and relative humidity illustrates the subsidence driven 
drying and the strong moistening that results from ascent over the region of maximum latent heat flux.

@@

Despite a fairly regular distribution of precipitation around the SST maximum for experiments with increasing resolution, the 
surface enthalpy flux reveals large differences in the symmetry.  Over the SST maximum, E25 has a surface enthalpy flux that is 
60 W/m2 larger than that of the E1 experiment, and the E1 experiment has a very irregular pattern of enthalpy flux in the middle
half of the domain.  It appears that these differences in magnitude and regularity are due to the differing wind speeds at the 
lowest atmospheric level.  It is also worth noting that the E1 experiment has a stronger domain mean u wind shear compared 
with the E2 and E25 experiments.  Over the warm patch of SST, the surface enthalpy flux is decreased by the LWCRE for E25, 
E2, and E1 despite these experiments having stronger low-level winds.  This decreased enthalpy flux also corresponds to 
an increased amount of specific humidity in the lowest atmospheric model level.        
%%%%%
% section 3 of results

@@

@@

Figure \ref{fig:cf_tdtlw} shows E2 to have the largest (about 17\%) upper level mean cloud
fraction in the subsidence region, with the E1 experiment having the next largest cloud fraction (10\%), followed by E25, P25, and 
P100 (5\%).   Less upper level clouds will allow more radiative cooling of the atmosphere and require a correspondingly larger
amount of condensational heating and precipitation.

@@

\begin{figure}
  \centering
       %\includegraphics[width=0.96\columnwidth]{walkerfigs/hist_6pan_25vs2vs1_lwoff_lwon.eps}
       \includegraphics[width=0.96\columnwidth]{walkerfigs/hist_3pan_25vs2vs1_lwoff_lwon_draft.eps}
  \caption{Normalized probability distributions of the location of the maximum value of precipitation for experiments with the 
  longwave CRE on.  All panels show cases with parameterized convection, the top panels are computed for the experiments 
  having a domain width of 4000 km while the bottom panels are for experiments having a domain width of 16000 km.  The 
  black curves show a Gaussian function with a standard deviation computed for each experiment and an expected location 
  in the center of the domain where the SST is warmest. }
\end{figure}

>>>>

% What are climate modelers doing to reduce uncertainties about cloud processes?
The representation of clouds in climate models is the most stubborn contributor to uncertainty of the climate response to increased 
greenhouse gases.  A few of the reasons clouds are difficult to represent in models are imprecise definitions of clouds, an existential 
dependence on trace amounts of condensed water, the importance of microphysical processes at scales less than the grid-scale, 
and the close coupling between atmospheric radiative fluxes, clouds, and the large-scale dynamical circulations of Earth.  To elucidate this 
coupling between clouds, radiation, and the large-scale circulations of the Earth, we simulate an idealized equatorial tropical Pacific.   In 
the western Pacific warm sea surface temperatures prevail beneath rising atmospheric motion and deep convective storms.   In contrast, 
the eastern tropical Pacific is characterized by cooler sea surface temperatures and low-level, non-precipitating clouds with a dry 
atmosphere above the clouds.  These regions are connected by a large-scale circulation.  We analyze this scenario in a high-resolution 
and coarse resolution simulation and infer how deep convective heating in the west Pacific can influence the atmospheric cooling 
thousands of kilometers away in the east Pacific.  This cooling is a key ingredient in the determination of the Earth's sensitivity to 
changes in greenhouse gas concentrations.

@@

\textbf{Paragraphs of Introduction}
\begin{itemize}
  \item{Citation of Walker, Bjerness, and Johnson et al. papers to set the stage}
  \item{The Walker Circulation as way of imposing realistic large-scale circulations, as opposed to spontaneously generated large-scale
  circulations.  Previous studies of RCE have proven to be a 
  useful tool with which to explore the basic physics of climate and climate sensitivity, but they lack a large-scale circulation.}
  \item{On the importance of reducing component, or sub-grid scale uncertainty through the use of both parameterized and explicit 
  convection.  This paragraph could also cite papers which have helped to validate RCE in general, and the channel RCE configuration
  in particular.}
\end{itemize}

@@

We perform a series of sensitivity experiments that highlight the different ways in which these experiments can equilibrate.  To 
highlight both the robust features of the simulations and the sensitive nature of the distribution of the precipitation maximum 
we test the impact of convective parameterization, 
longwave radiation interactions with clouds, domain size, 
and the resolution, or grid-spacing.    Muller and Held, 2012 found that self-aggregation was sensitive to both domain size and resolution
because of the sensitivity of the low-level clouds to the longwave radiative cooling.

@@

\begin{table}
\begin{center}
\caption{A few specifications of the simulations used most heavily in this study.  In the Convection
column, `prm'  indicates that convection is parameterized and `exp' indicates convection is explicity.  
The precipitation values in parenthesis are for the experiments in which LWCRE is off.}
    \begin{tabular}{*{6}{c}}
    \hline
    \hline
    \\
 Name & Grid Spacing $(\mathrm{km})$ & Domain $ (\mathrm{km^2}) $& Length $(\mathrm{months}) $ & Convection & 
 $\overline{\rm{P}} (\mathrm{mm/day})$    \\ \hline
  P100L &  100          &   800 $\times$ 16000    &  60              & prm           &  4.1 (3.5)          \\ 
    \\
  P100 &  100                & 800 $\times$ 4000     & 60            & prm         & 3.9 (3.7)               \\  
    \\
  P25L &  25             & 200 $\times$ 16000      & 60             & prm          & 4.0 (3.8)                \\  
    \\
  P25  &  25             & 200 $\times$ 4000      & 60             & prm           & 3.8 (3.7)               \\  
    \\
 E25  &   25          & 200 $\times$ 4000      & 60             & exp          & 3.7 (3.5)               \\  
    \\
 E2   &   2          & 100 $\times$ 4000      & 6             & exp          & 3.1 (3.4)                \\ 
    \\
 E1   &   1          & 10 $\times$ 4000      & 6             & exp          & 3.3 (3.7)                \\  \hline

\begin{table}
\begin{center}
\caption{Domain mean precipitation ($\overline{\rm{P}}$) and outgoing longwave radiation 
($\overline{\rm{OLR}}$).
The values in parenthesis are for the experiments in which LWCRE is off.}
    \begin{tabular}{*{3}{c}}
    \hline
    \hline
    \\
 Name &   $\overline{\rm{P}} (\mathrm{mm/day})$ & $\overline{\rm{OLR}} (\rm{W \, m^-2})$   \\ \hline
  P100L   &  4.1 (3.5)   &  283.1 (286.9)          \\ 
    \\
  P100 &   3.9 (3.7)   &  283.2 (296.4)             \\  
    \\
  P25L &   4.0 (3.8)  &  281.2 (290.7)             \\  
    \\
  P25  &   3.8 (3.7)    & 282.9 (293.6)             \\  
    \\
 E25  & 3.7 (3.5)    &   271.9 (286.8)               \\  
    \\
 E2   &  3.1 (3.4)   &  266.2 (285.5)               \\ 
    \\
 E1   &  3.3 (3.7)   &  269.3 (289.2)               \\  \hline

@@

Superposing the circulation and relative humidity illustrates the subsidence driven 
drying and the strong moistening that results from ascent over the region of maximum latent heat flux.  
Figure \ref{fig:rh_psi_P25vsE25} shows two distinct circulation cells with one below, and one above $500 \rm{hPa}$.  Deep-convective activity dominates the 
region over the warm pool, shallow convection is common over a wider range of SSTs, and for the most part stratocumulus 
clouds are absent.  When the coupling between the circulation and clouds is broken, by turning off the longwave CRE, the atmospheric 
state is much more symmetric about the maximum SST.  
We also find large differences among the experiments in the domain mean precipitation, condensate/clouds, and 
circulation (see table 1).  In particular
the experiments with parameterized convection are much more erratic.

One of the most prominent features of our GCM-like (in particular, P25) simulations is an asymmetry in both the time-dependent and 
steady-state solutions (Figures \ref{fig:rh_psi_P25vsE25},  \ref{fig:precip_vertvel}, \ref{fig:domdep} and  \ref{fig:conv_vs_ls}).  The steady-state precipitation maximum is located  not over the warmest SST but is
shifted to slightly cooler temperatures.  The asymmetry is apparent in the vertical velocity, mass circulation, relative humidity, 
specific humidity, and radiative heating.  
For most of these cases the asymmetry appears as an aversion of the precipitation maxima to reside over the SST maximum.  
For the P25 case shown in Figure \ref{fig:rh_psi_P25vsE25} a strong ($1 \, \rm{m \, s^-1}$) domain
mean shear develops above about 500 \textit{m} which shifts the precipitation and circulation off center for years at a time.
When the convective parameterization
is turned off (E25), the overturning circulation becomes weaker and broader (w, and psi), and the precipitation, cloud fields, and 
circulation all consistently reside over the SST maximum.   
While the parameterized convection plays a large role in driving this asymmetry it appears to be not entirely a 
result of the convective parameterization, but also due to an interaction between the convective 
parameterization and the LW CRE.   The degree to which this asymmetry influences the comparison with other experiments is 
unclear.    Note: check the wind shear for P25large, P100small, and P100large (E25 has less shear than E1 or E2).

Complex patterns of precipitation over a fixed sinusoidal or Gaussian SST 
distribution have been noted many times in previous literature
(Bretherton et al. 2006, Kuang and Wofsy, Ming Zhao (personal communication, see Jeevanjee et al. 2018)) but the irregularities have 
tended to be symmetric about the SST maximum.  This is broadly consistent with our simulations when the convection is 
entirely explicit (discussed further in section 4).

@@

@@

Now talk about the organizing influence of LWCRE...

@@

Because of the close connection between the domain mean value of precipitation, atmospheric condensational heating, and 
the total radiative cooling the time evolution of the precipitation serves as a useful measure of whether a model has 
reached statistical equilibrium.  Figure \ref{fig:precip_dom_mn} demonstrates that after about 30 days, the P25, E2, and E1
simulations begin to oscillate about a mean state that is roughly constant over periods greater than about 30 days. 
The large oscillations in domain mean precipitation are similar to those noted in previous studies 
(Silvers et al., 2016; Patrizio and Randall, 2019).   The experiments with a higher resolution have a lower value of domain
mean precipitation, with the E2 simulation having the lowest (3.1 mm/day).  This can be understood as a consequence
of the differences in upper level cloud fraction and will be discussed further in a later section.

@@

\section{The Influence of Domain Size on the LWCRE and the LS Parameterized Precipitation: Change Title?}

@@

When designing numerical experiments the size of the domain and the grid-spacing that determines model resolution are two of the most 
fundamental choices that must be made.   One the goals of this study was to explore the results of changing grid-spacing over a wide 
range of values.  To simplify the analysis we have in most cases chosen to keep the 
long horizontal dimension fixed at $4000\, \rm{km}$.  However, previous studies (e.g. Bretherton et al., 2006; Muller and Held, 2012; Jeevanjee and Romps, 2013) have 
documented sensitivities to domain size.   The analysis of the previous, and of the next section focus on results from experiments using
a domain with a long domain length of $4000 \, \rm{km}$.  However we do find sensitivity to the domain size and thought it to be
sufficiently interesting to report.

Examining the evolution in time of the precipitation field reveals important characteristics of these simulations, and 
illustrates how much the distribution can vary as a function of domain size, parameterization of convection, and the effect of the 
longwave radiation due to clouds. 
Shown in Figures {\ref{fig:domdep}} and {\ref{fig:domdep_lwoff}} are Hovmoller plots of precipitation after 
averaging along the short horizontal dimension.  The four panels show simulations with a grid-
spacing of $25\, \rm{km}$ on a grid with a horizontal width of $4000\, \rm{km}$ (far right) and $16000\, \rm{km}$ (middle right), and 
simulations with a grid-spacing of $100\, \rm{km}$ on a grid with a horizontal width of $4000\, \rm{km}$ (middle left) and 
$16000\, \rm{km}$ (far left).  Figure  {\ref{fig:domdep_lwoff}} shows the equivalent simulations with LW CRE off.    
 At all resolutions the Hovmoller plots show that the longwave CRE acts to concentrate the 
precipitation over a smaller geographic extent.   This is consistent with the idea that CREs strengthen an overturning circulation and
narrow the region of deep convection (see Popp and Silvers, 2017, HH, Dixit et al. 2019.  etc.).
The structure of the precipitation changes more as a function of domain size than it does as a function of grid-spacing. 
On the large domains, the difference between experiments with and without LW CRE is extreme.  In contrast to the control 
experiments in Figure {\ref{fig:domdep}} which all show a narrow region of strong precipitation which meanders within 
500 km of the SST maximum at the center of the domain,  
the experiments without the LW CRE basically have no region with consistently strong precipitaiton.  Instead, 
smaller cells or lines of precipitation develop within an area that is roughly 8000 km wide.

@@

Smaller domain lengths were found to have a more focused ascent region and larger precipitation rates in Bretherton et al., 2006.
They also found less low-level clouds over the colder SSTs in the small domains, relative to a control domain size that was 4x
larger.   We have found larger precipitation rates for the larger domains.  Does our larger domain ascent region change?

@@

Although the domain mean total precipitation is constrained by the radiative cooling of the atmosphere, the amount of 
precipitation that is produced by the convection scheme or the large-scale cloud scheme is not constrained. 
The distribution of convective and large-scale precipitation can often give a hint of how the condensational heating
in a GCM is being distributed among the parameterizations.  Precipitation from each of these two components is
shown in Figure \ref{fig:conv_vs_ls}.  From this we can see that in the regions of large-scale ascent, most of the 
precipitation derives from the large-scale cloud scheme.  We also see that the LW CRE strongly influences the 
large-scale precipitation but has a much weaker impact on the convective precipitation.  With the single exception 
of the P100 LW CRE off experiment, the convective precipitation
produces relatively little of the total precipitation for both resolutions and on both domains.

%We hypothesize than an important difference between our simulations and previous studies which use GCMs
%is that although the deep and shallow convection is parameterized
%with the GFDL double-plume parameterization, the fraction of precipitation that is produced by the convective parameterization 
%is much smaller in our simulations than it is for many GCM-type experiments with parameterized convection.   At 25km grid-spacing, 
%the majority of condensational heating occurs through the large-scale cloud  scheme (Figure \ref{fig:conv_vs_ls}).
%%Overall, the fraction of precipitation that is 
%%due to the large-scale scheme is roughly 40\%, and in many places it is actually greater than the amount of precipitation 
%%produced by the convection scheme.   
%%In comparison, many GCMs, have between 70\%-90\% of their precipitation produced by 
%%the convective parameterization.  
%There are further surprises in the fraction of precipitation that is large-scale.  On our small 
%domains, with a bowling alley domain length of 4000 km, the simulations with 100km grid-spacing have values of $f_{ls}$ 
%of $75-80\%$.  Often, at GCM-like resolutions such as 100km the convective precipitation would be expected to control much of the
%precipitation rather than the large-scale parameterization.  However, when the domain length is increased to 16000 km
%$f_{ls}$ decreases to a range of $0-70\%$ that depends strongly on location.  Over the warmest region of the surface the 
%precipitation is still dominated by the large-scale scheme.   The large domain without the longwave CRE is the 
%only configuration in which the convective precipitation is dominant.    If the default, or baseline, configuration of a GCM is built 
%with the expectation that the convective parameterization scheme will dominate the production of tropical precipitation, then 
%perhaps a configuration that minimizes the convective precipitation will result in a balance between the parameterization 
%schemes that is less intuitive, such as an asymmetric response to symmetric forcing.

Below is a list summarizing several of the changes that occur when the LW CRE is off.  The points which have not already been discussed will be later in the paper.   Turning the LW CRE interactions off tends to:

\begin{itemize}
  \item decrease the strength of the overturning circulation and the domain mean precipitation (P100,P25,E25)
  \item increase the midtropospheric RH by $5-15 \% $
  \item eliminate the difference of liquid condensate at 900 hPa between explicit and parameterized simulations.
  \item weaken the horizontally oriented low-level circulations.   
  \item spread out in geographic space the precipitation maxima
  \item homogenize the surface enthalpy flux over the warm pool, primarily through the latent heat flux
  \item decrease both the LTS and the EIS, over both the warm and cold regions
  \item decrease the lowest level $q, \theta_e,$ temperature $T$ and virtual temperature $T_v$.  
  \item eliminate large differences in the evaporation and lowest level temperature in the center of the domain between 
           the 1 and 2 \textit{km} simulations. 
  \item increase evaporation in the 1 and 2 \textit{km} cases, decrease it in the 25 \textit{km} case.
  \item amplify the larger values of ice condensate aloft for the explicit simulations.  
\end{itemize}

@@

It is clear that the overturning circulation in the tropical Pacific encompasses dynamical motions at a host of scales and that the 
different cloud types interact with the circulation on scales ranging from meters to thousands of km's.  
The tropical Pacific also serves as an ideal location to study interactions between clouds and the 
circulation because it includes strong overturning circulations, abundant shallow cumulus, congestus, and cumulonimbus 
clouds (Johnson et al. 1999), as well as stratocumulus cloud decks along the eastern extremities of the basin.
Can we hope to simulate all of this in a single idealized modeling context?    
Our initial motivation for using the GFDL AM4 model on a doubly-periodic domain was to simulate this region 
%that is rich in phenomena 
in the context of both a GCM and a CRM in the hopes that resolving more of the turbulent motions and circulations
would help us to better understand the physics and the mechanisms which are at work in the cloud-circulation
interactions of the tropical Pacific.    
The dynamical core of the current generation GFDL models is ideally suited to this task because it can easily 
be used over a wide range of resolutions and can solve either the hydrostatic or nonhydrostatic 
governing equations.  In this section we focus on simulations with grid-spacing of 1km, 2km, and 25km.   
A glance at Figure \ref{fig:hov_4mods_6mn} shows notable differences in the precipitation structure that result from the 
overturning circulation at different resolutions.  Shown are 180 days of precipitation from the P25, E25, E2, and 
E1 simulations.  As the 
resolution increases towards $1 \, \rm{km}$ the distribution of precipitation becomes more regular and consistently 
centered over the SST maxima in the middle of the domain.  Both the P25 and E25 simulations show more 
variability at later 
times, compared to these first 180 days.  The simulations with explicit convection at resolutions typical of 
cloud-resolving 
models (E2, E1) show little aversion to the precipitation maximum occurring over the maximum in SST and relative to 
the P100, P25 and E25 simulations 
they are able to maintain a smoother distribution of  precipitation values over a broader range of SST values.

@@

@@

Despite a fairly regular distribution of precipitation around the SST maximum for experiments with increasing resolution, the 
surface enthalpy flux reveals large differences in the symmetry.   Figure \ref{fig:enthalpy} shows the surface enthalpy flux, 
the equivalent potential temperature, and the u-component wind field for E1,E2, and E25 both with (thick lines) and 
without (thin lines) the LW CRE.  
Over the SST maximum, E25 has a surface enthalpy flux that is 
60 W/m2 larger than that of the E1 experiment, and the E1 experiment has a very irregular pattern of enthalpy flux in the middle
half of the domain.  It appears that these differences in magnitude and regularity are due to the differing wind speeds among 
the three experiments.  The asymmetry in the enthalpy flux is also apparent in the lowest model level of the atmosphere
and the E1 experiment also has the largest domain mean wind shear of the three models.  This is consistent with the 
strong asymmetries of the P25 experiment being connected to a large domain mean wind shear in that experiment.  
It is also worth noting that the E1 experiment has a stronger domain mean u wind shear compared 
with the E2 and E25 experiments.  Figure \ref{fig:enthalpy} also shows that over the warm patch of SST, the surface 
enthalpy flux is decreased by the LWCRE for E25, E2, and E1 despite these experiments having stronger low-level winds.  
This decreased enthalpy flux also corresponds to an increased amount of specific humidity in the lowest atmospheric 
model level.        
%%%%%
% section 3 of results

@@

%\begin{figure}
%  \centering
%      \includegraphics[width=0.6\columnwidth]{walkerfigs/Precip_DMN_4models_1yr.eps}
%  \caption{Domain mean precipitation as a function of time.  All data has been smoothed with a running mean filter.  
%  Thick lines are the control experiments, thin lines have no LW CRE.  All data is shown for the E1 and E2 cases
%  while the P25 and P100 cases show only the first out of five years of the time series.}
%    \label{fig:precip_dom_mn}
%\end{figure}

@@

Figure \ref{fig:cf_tdtlw} shows E2 to have the largest (about 17\%) upper level mean cloud
fraction in the subsidence region, with the E1 experiment having the next largest cloud fraction (10\%), followed by 
P100, E25, and P25 (3-5\%).   Less upper level clouds will allow more radiative cooling of the atmosphere and 
require a correspondingly larger amount of condensational heating and precipitation.

@@

\begin{figure}
  \centering
       %\includegraphics[width=0.96\columnwidth]{walkerfigs/hist_6pan_25vs2vs1_lwoff_lwon.eps}
       \includegraphics[width=0.96\columnwidth]{walkerfigs/hist_3pan_25vs2vs1_lwoff_lwon_draft2.eps}
  \caption{Normalized probability distributions of the location of the maximum value of precipitation for experiments with the 
  longwave CRE on.  All panels show cases with parameterized convection, the top panels are computed for the experiments 
  having a domain width of 4000 km while the bottom panels are for experiments having a domain width of 16000 km.  The 
  black curves show a Gaussian function with a standard deviation computed for each experiment and an expected location 
  in the center of the domain where the SST is warmest. }
\end{figure}

An interesting fact that emerges from the domain mean values of precipitation (see Table 1) 
is that the sign of the response to LWCRE is not the same across the different grid-spacings.
When clouds are not allowed to interact with the longwave radiation the atmosphere can cool 
much more efficiently, as evidenced by larger values of OLR for all experiments when the LWCRE is
off.  Typically, the atmospheric cooling can be thought of as a proxy for the mean precipitation 
because the cooling is primarily offset by condensational heating.  This is clearly not the case 
for the E25, P25, and P100 cases for which the domain mean precipitation rates decrease despite
an increased amount of atmospheric cooling.  The implication is that the requisite atmospheric 
heating is coming from a process other than condensation.  The solution lies in the energy 
budget of the surface.  A fixed SST means that in all cases the upward flux of longwave energy
is constant, the upward flux of sensible heat flux will be mostly fixed (barring variations in surface
wind), and changes in downward shortwave flux will not be able to warm the surface.  However, 
the dramatic decrease of low-level clouds for the E25, P25, and P100 experiments when LWCRE is
